---
title: "Silverio_Vasquez_Hw3"
author: "Silverio Vasquez"
date: "September 17, 2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 3


```{r Q3}
library(stringr)

raw.data <- "555-1239Moe Szyslak(636) 555-0113Burns, C. Montgomery555-6542Rev. Timothy Lovejoy555 8904Ned Flanders636-555-3226Simpson, Homer5553642Dr. Julius Hibbert"

name <- unlist(str_extract_all(raw.data, "[[:alpha:]., ]{2,}"))

phone <- unlist(str_extract_all(raw.data, "\\(?(\\d{3})?\\)?(-|)?\\d{3}(-|)?\\d{4}"))
```

1 - convert to first_name and last_name
```{r Q3_1}
name_wo_comma <- unlist(name[!str_detect(name,",")])

name_wo_pref <- name
for (pref in c("Mr.","Ms.","Mrs.","Dr.","Rev.")){
        name_wo_pref <- unlist(str_trim(str_replace_all(name_wo_pref,pref,"")))
}
name_w_comma <- unlist(str_split(name_wo_pref[str_detect(name_wo_pref,",")],","))

name1 <- str_c(name_w_comma[2],name_w_comma[1],sep="_")
name2 <- str_c(name_w_comma[4],name_w_comma[3],sep="_")

name_wo_comma <- unlist(name_wo_pref[!str_detect(name_wo_pref,",")])
name_wo_comma <- unlist(str_replace_all(name_wo_comma," ","_"))

ex1_ans <- c(name_wo_comma,str_trim(name1),str_trim(name2))
print(ex1_ans)
```


2 - logical vector for titles
```{r Q3_2}
ex2_ans <- unlist(str_detect(name,"Dr.")) | unlist(str_detect(name,"Rev."))
print(ex2_ans)
```

3 - logical vector for a second name. 

Assumption - first name and last name will only be separated by one space. So, if there's a middle name, then there will be more than one space.
```{r Q3_3}
ex3_ans <- as.logical(str_count(ex1_ans,"\\s"))
print(ex3_ans)
```

## Question 4
a. The pattern this describes is any number before the end of a sentence signaled by a "\$". For example "25$" will yield "25".

b. The pattern this describes is any word that's four lower-case letters or less. For example, "Silo vase Case base1" will yield only "vase"

c. This doesn't describe any pattern. first ".*" will repeat the entire string inputted but "?" will turn everything into empty quotes. 

d. This describes a very specific pattern where there needs to be at least two digits followed by "/" then only two digits, followed by "/" then followed by at least four digits.

e. Not really sure what this pattern is because I tried it on different strings and kept getting nothing.

## Question 9
```{r Q9}
q9 <- paste0("clcopCow1zmstc0d87wnkig7OvdicpNuggvhryn92Gjuwczi8hqrfpRxs5Aj5dwpn0Tanw","Uwisdij7Lj8kpf03AT5Idr3coc0bt7yczjatOaootj55t3Nj3ne6c4Sfek.r1w1YwwojigO",
"d6vrfUrbz2.2bkAnbhzgv4R9i05zEcrop.wAgnb.SqoU65fPa1otfb7wEm24k6t3sR9zqe5",
"fy89n6Nd5t9kc4fE905gmc4Rgxo5nhDk!gr")
q9 <- str_conv(q9,"UTF-8")
q9_msg <- unlist(str_extract_all(q9,"[[:upper:][:punct:]]"))
print(q9_msg)
```
Question 9 --> "CONGRATULATIONS. YOU ARE. A. SUPERNERD!"