---
title: "SVasquez_HW4"
author: "Silverio Vasquez"
date: "September 29, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(curl)
```

## Get Data and Tidy Up

Below code brings in a "wide" dataset from a CSV file in Github, drops row if there's an 'na' in the second column, and fills empty cells in column 1 with prior values. 

```{r import}
flights <- read_csv(curl("https://raw.githubusercontent.com/sjv1030/Data607-hw/master/flights.csv"))

head(flights)
flights <- flights %>% drop_na(X2) %>% fill(X1)
```

Below code "gathers" the data by city and then spreads it by status ("delayed"" or "on_time""). Mutate adds a total column ("tot"") that's used to calculate delayed flights as a percent of total flights (column "pct"). All of this converts the data set from a "wide" format to a "long" format.

```{r tidyup}
df <- flights %>% gather('city','n',3:7) %>% spread(X2,n)
names(df) <- c('airline','city','delayed','on_time')

df <- df %>% mutate(tot = delayed + on_time)
df <- df %>% mutate(pct = round(delayed/tot,4) * 100)

head(df,10)
```

## Interesting Factoids

Which airline had the largest percent of delays and to which city?


```{r f1}
df[which.max(df$pct),]
```


Which airline had the lowest percent of delays and to which city?


```{r f2}
df[which.min(df$pct),]
```

Show average percent of delays by airline:


```{r f3}
f3 <- df %>% group_by(airline) %>% summarise(avg=mean(pct)) %>% arrange(desc(avg))
ggplot(f3,aes(airline,avg))+ggtitle('Average Percent of Delays by Airline')+
        geom_col(fill="#0072B2",colour="black")+theme_minimal()+xlab('')+ylab('')
```

Show max percent of delays by airline:


```{r f4}
f4 <- df %>% group_by(airline) %>% summarise(high=max(pct)) %>% arrange(desc(high))
ggplot(f4,aes(airline,high))+ggtitle('Max Percent of Delays by Airline')+
        geom_col(fill="#0072B2",colour="black")+theme_minimal()+xlab('')+ylab('')
```

Show max percent of delays by city:


```{r f5}
f5 <- df %>% group_by(city) %>% summarise(high=max(pct)) %>% arrange(desc(high))
ggplot(f5,aes(city,high))+ggtitle('Max Percent of Delays by City')+
        geom_col(fill="#0072B2",colour="black")+theme_minimal()+xlab('')+ylab('')
```


Given the data, we can conclude that Phoenix is the best airport to have a layover in and San Fran is the worst. Additionally, it looks like Alaska airlines is a lot more reliable than AM West.